---
author: "Khunanon Chanasongkhram"
date: "24/01/2023"
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# DisVar: An R library for identifying variants associated with diseases using personal genetic information

<!-- badges: start -->
<!-- badges: end -->
DisVar is an R package that finds diseases from a VCF file by comparing the variants in the VCF file with those in various databases. The package currently supports five databases: GWASdb, GRASP, GWAS Catalog, GAD, and Johnson and O'Donnell.

## Features

- Finds diseases from a VCF file by comparing the variants in the VCF file with those in various databases.

- Supports five databases: GWASdb, GRASP, GWAS Catalog, GAD, and Johnson and O'Donnell.

## Installation

You can install the development version of DisVar from [GitHub](https://github.com/) to install the package, you can use the `devtools` package and run the following command:

``` r
# install.packages("devtools")
devtools::install_github("Khunanon-Chanasongkhram/DisVar")
```
## Dependencies
The package depends on the following packages:

- `sqldf`
- `data.table`

## Usage

The main function in the package is `DisVar()`, which takes two arguments:

- `file`: the name of the VCF file.

- `skip`: the number of lines to skip at the beginning of the VCF file. (default: 0)

The function returns a data frame containing variant information from various databases.

### Examples

Here's an example of how to use the function:

```{r example}
library(DisVar)
DisVar("file_name.vcf", skip = 28)
```

You will get this output file: `file_name`_diseases_output.txt

When you test the program with the test file (`file_name.vcf`), the results are shown in the following table.

| Disease                                                    | Chr | Position | Variant_id | Allele sample | Allele DB | Confident/P-value | DB           |
|------------------------------------------------------------|-----|----------|------------|---------------|-----------|-------------------|--------------|
| Breast cancer                                              |   3 |  4700592 | rs6762644  | A>G           | A>G       |          2.00E-12 | GWASdb       |
|                                                            |   3 |  4700592 | rs6762644  | A>G           | NA        |          2.20E-12 | GRASP        |
| Breast cancer (Estrogen receptor positive   breast cancer) |   3 |  4700592 | rs6762644  | A>G           | NA        |          1.40E-08 | GRASP        |
| Breast cancer   (invasive breast cancer)                   |   3 |  4700592 | rs6762644  | A>G           | NA        |          1.20E-09 | GRASP        |
|                                                            |   3 |  4700592 | rs6762644  | A>G           | NA        |          2.00E-12 | GWAS Catalog |
|                                                            |   3 |  4700592 | rs6762644  | A>G           | NA        |          4.00E-18 | GWAS Catalog |
| Breast_cancer                                              |   3 |  4700592 | rs6762644  | A>G           | NA        |          9.00E-12 | GWAS Catalog |


## Additional Resources

- The VCF format is a standard file format for storing genetic variation data, and it is widely used in bioinformatics research. More information on the VCF format can be found at the [VCF specification page](http://samtools.github.io/hts-specs/VCFv4.3.pdf)

## Limitations and Known Issues

- The package is limited to searching for variants in specific databases, and may not include all known disease-associated variants.

- The package uses external databases, which are subject to change and may contain errors.

- The package is still under development and may have bugs. If you encounter any issues, please report them on the GitHub repository.

- The package may take a long time to process large amounts of data.

[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/DisVar)](https://cran.r-project.org/package=DisVar)


